Fast, Scalable Phrase-Based SMT Decoding

### 2016-02-08 acl.2016/acl2016.tex
acl paper

We present in this paper a re-assessment of the major components of phrase-based decoding and decoder implementation with particular emphasis on speed and scalability to multicore machines. The result is a drop-in replacement for the Moses decoder which is up to fifteen times faster and scales almost linearly with the number of cores. Furthermore, the decoder makes less search errors than the current Moses decoder.

@@

~\newcite{Heafield-mtplz} is perhaps closest in intent to this work. This takes a wholistic approach to decoding, describing a novel decoding algorithm which is fis focused on better decoding speed. It also describes a number of implementation details for faster decoding. However, the decoding algorithm is only able to incorporate one stateful feature function which precludes some of the useful decoding configurations which contains multiple stateful feature functions. It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment where phrase-table has not be filtered with a know test set for any realistic size phrase-table. Neither did this paper analyze the scalability of their work to multicore servers.

The rest of the paper will be broken up into the following sections. Next, we will describe the phrase-based model and the major implementation components, with particular emphasis on decoding time shortcomings. We will then describe modifications to improve decoding speed and present results. We conclude in the last section discuss suggested improvements and future work.

@@

The first three feature functions frequently trained on data and require the feature to read the model from files. The other feature functions do not require model files. We shall investigate the first two feature functions for optimization.

@@

Overall though, the scalability is remarkably good, the decoder is able to make full use of all real and virtual cores. When all cores are saturated, the decoder is 16 times faster than single-threaded decoding, on this 16 real core server. It is also 4.5 times faster than Moses with a single-thread and 14.5 faster when all cores are used.

@@

In future, we shall work on exploring the underlying reasons for the scalability issues in Moses to get a better understanding where potential performance issues can arise. We shall also investigate other major components of the decoding algorithm, particulaly the language model which has not been touched in this paper.

>>>>

We present in this paper a re-assessment of the major components of phrase-based decoding and decoder implementation with particular emphasis on speed and scalability to multicore machines. The result is a drop-in replacement for the Moses decoder which is up to fifteen times faster and scales almost linearly with the number of cores.

@@

~\newcite{Heafield-mtplz} is perhaps closest in intent to this work. This takes a holistic approach to decoding, describing a novel decoding algorithm which is fis focused on better decoding speed. It also describes a number of implementation details for faster decoding. However, the decoding algorithm is only able to incorporate one stateful feature function which precludes some of the useful decoding configurations which contains multiple stateful feature functions. It does not include a load-on-demand phrase table, therefore, cannot be used in a commercial environment where phrase-table has not be filtered with a know test set for any realistic size phrase-table. Neither did this paper analyze the scalability of their work to multicore servers.

The rest of the paper will be broken up into the following sections. Next, we will describe the phrase-based model and the major implementation components, with particular emphasis on decoding speed. We will then describe modifications to improve decoding speed and present results. We conclude in the last section and discuss future work.

@@

The first three feature functions frequently trained on data and require the feature to read the model from files. The other feature functions are simple and do not leave much room for optimization. The language model consumes a large part of decoding time, and it is for this reason that it has already been heavily optmized. We shall therefore investigate the first two feature functions for optimization.

@@

Overall though, the scalability is remarkably good, the decoder is able to make full use of all real and virtual cores. When all cores are saturated, the decoder is 16 times faster than single-threaded decoding, on this 16 real core server. It is also 4.5 times faster than Moses with a single-thread and 9.6 faster when all cores are used.

@@

In future, we shall investigate other major components of the decoding algorithm, particulaly the language model which has not been touched in this paper. We shall also explore the underlying reasons for the scalability issues in Moses to get a better understanding where potential performance issues can arise. This has application to other algorithms beside MT decoding.
